---
title: "R Data Structure"
subtitle: "Solution Exercises"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 3
    collapsed: false
    css: ["assets/css/custom_report.css"]
linestretch: 1.5
number_sections: true
link-citations: true 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```


# Operators

1. Compute $\frac{(45+21)^3+\frac{3}{4}}{\sqrt{32-\frac{12}{17}}}$.
  <details><summary>See solution</summary>
    ```{r}
    ((45 + 21)^3 + 3/4)/(sqrt(32 - 12/17))
    ```
  </details>
1. Compute $\frac{\sqrt{7-\pi}}{3\ (45-34)}$.
  <details><summary>See solution</summary>
    ```{r}
    (sqrt(7) - pi)/(3 * (45 - 34))
    ```
  </details>
1. Define a proposition to evaluate if a number is even.
  <details><summary>See solution</summary>
    ```{r}
    value <- 3
    value %% 2 == 0
    
    value <- 4
    value %% 2 == 0
    ```
  </details>
1. Define a proposition to evaluate if a number is between -4 and -2 or between 2 and 4.
  <details><summary>See solution</summary>
    ```{r}
    value <- 3
    abs(value) >= 2 & abs(value) <= 4
    ```
  </details>
1. Compare the returned results by `4 ^ 3 %in% c(2,3,4)` and `4 * 3 %in% c(2,3,4)`.
  <details><summary>See solution</summary>
    ```{r}
    4 ^ 3 %in% c(2,3,4)
    
    4 * 3 %in% c(2,3,4)
    ```
  </details>

# Vectors

### Creating {-}

1. Create `x` with all even numbers between 1 e 25 .
  <details><summary>See solution</summary>
    ```{r}
    x <- seq(from = 2, to = 25, by = 2)
    x
    ```
  </details>
1. Create `y` with the first 10 multiples of 7 starting from 14.
  <details><summary>See solution</summary>
    ```{r}
    y <- seq(from = 14, by = 7, length.out = 10)
    y
    ```
  </details>
1. Create `a` with the letters `"A"`,`"B"` e `"C"` repeated in the same order 4 times.
  <details><summary>See solution</summary>
    ```{r}
    a <- rep(c("A", "B", "C"), times = 4)
    a
    ```
  </details>
1. Create `b` with the letters `"A"`,`"B"` e `"C"` each one repeated 4 times.
  <details><summary>See solution</summary>
    ```{r}
    b <- rep(c("A", "B", "C"), each = 4)
    b
    ```
  </details>


### Selecting {-}

1. From `x`, select the values 8, 12, and 16.
  <details><summary>See solution</summary>
    ```{r}
    x[c(4, 6, 8)]
    ```
  </details>
1. From `y`, select values smaller than 36 or grater than 54.
  <details><summary>See solution</summary>
    ```{r}
    y[y < 36 | y > 54]
    ```
  </details>
1. Create `c` from `a` removing all `"B"` elements.
  <details><summary>See solution</summary>
    ```{r}
    c <- a[a != "B"]
    c
    ```
  </details>
1. Crate `d` from `b` substituting all `"B"` with `"D"`.
  <details><summary>See solution</summary>
    ```{r}
    d <- b
    d[d == "B"] <- "D"
    d
    ```
  </details>

# Factors


1. Create the factor `gender` as follows,
    ```{r, echo=FALSE}
    factor(c(rep(c("M","F"),3),"F","F","M"))
    ```
  <details><summary>See solution</summary>
    ```{r}
    gender <- factor(c("M", "F", "M", "F", "M", "F", "F", "F", "M"))
    gender
    ```
  </details>
1. Rename levels in `"Males"` and `"Females"`
  <details><summary>See solution</summary>
    ```{r}
    levels(gender) <- c("Females", "Males")
    gender
    ```
  </details>
1. Create the factor `size` as follows,
    ```{r, echo=TRUE, eval = FALSE}
    c("Small", "Large", "Medium", "X-Large", "X-Small")
    ```
  <details><summary>See solution</summary>
    ```{r}
    size <- factor(c("Small", "Large", "Medium", "X-Large", "X-Small"))
    size
    ```
  </details>
1. Make `size` an ordered factor, with appropriate label order.
  <details><summary>See solution</summary>
    ```{r}
    size <- factor(c("Small", "Large", "Medium", "X-Large", "X-Small"),
                   levels = c("X-Small", "Small", "Medium", "Large", "X-Large"),
                   ordered = TRUE)
    size
    ```
  </details>

# Matrices

### Creating {-}


1. Create the matrix `A` as follows, 

$$
\begin{matrix}
2 & 34 & 12 & 7\\
46 & 93 & 27 & 99\\
23  & 38 & 7 & 04
\end{matrix}
$$
  <details><summary>See solution</summary>
    ```{r}
    A <- matrix(c(2, 34, 12, 7, 46, 93, 27, 99, 23, 38, 7, 04),
                ncol = 4, byrow = TRUE)
    A
    ```
  </details>
1. Create the matrix `B`4X3 in which `"A"`,`"B"` e `"C"` are repeated at each row.
  <details><summary>See solution</summary>
    ```{r}
    B <-  matrix(c("A", "B", "C"), ncol = 3, nrow = 4, byrow = TRUE)
    B
    ```
  </details>

### Selecting {-}

1. From `A`, select the value 27.
  <details><summary>See solution</summary>
    ```{r}
    A[2, 3]
    ```
  </details>
1. From `B`, select the elements between the second and forth row, second and third column.
  <details><summary>See solution</summary>
    ```{r}
    B[2:4, c(2,3)]
    ```
  </details>
1. Form `A`, select only odd values.
  <details><summary>See solution</summary>
    ```{r}
    A[A %% 2 == 1]
    ```
  </details>
1. From `B`, select all values different from `"C"`.
  <details><summary>See solution</summary>
    ```{r}
    B[B != "C"]
    ```
  </details>

# Dataframe

### Creating {-}

1. Create the following dataframe

```{r, echo = FALSE}
my_data<-data.frame(Id = c("subj_1","subj_2","subj_3"),
                      age = c(21,23,19),
                      gender = c("F","M","F"),
                      item_1 = c(2,1,1),
                      item_2 = c(0,2,1),
                      item_3 = c(2,0,1))
my_data
```

  <details><summary>See solution</summary>
    ```{r}
    my_data <- data.frame(Id = c("subj_1","subj_2","subj_3"),
                          age = c(21,23,19),
                          gender = c("F","M","F"),
                          item_1 = c(2,1,1),
                          item_2 = c(0,2,1),
                          item_3 = c(2,0,1))
    my_data
    ```
  </details>
  
### Selecting {-}

1. Select data of `"subj_3"`.
  <details><summary>See solution</summary>
    ```{r}
    my_data[my_data$Id == "subj_3", ]
    ```
  </details>
1. Select only the answers to the items.
  <details><summary>See solution</summary>
    ```{r}
    my_data[, c("item_1", "item_2", "item_3")]
    ```
  </details>
1. Select `Id` and `gender` for subject with `1` at `item_1`.
  <details><summary>See solution</summary>
    ```{r}
    my_data[my_data$item_1 == 1, c("Id", "gender")]
    ```
  </details>

# Lists

### Creating {-}

1. Create a list with the elements form the previous exercises:
  - vector `x`
  - matrix `A`
  - dataframe
  <details><summary>See solution</summary>
    ```{r}
    my_list <- list(x = x,
                    A = A,
                    my_data = my_data)
    my_list
    ```
  </details>
  
### Selecting {-}

1. Select the vector and the matrix from the list.
  <details><summary>See solution</summary>
    ```{r}
    my_list[1:2]
    ```
  </details>
1. Extract the dataframe from the list.
  <details><summary>See solution</summary>
    ```{r}
    my_list[["my_data"]]
    ```
  </details>
1. Add the sting `"Hello World!"` to the list.
  <details><summary>See solution</summary>
    ```{r}
    my_list$string <- "Hello World!"
    my_list
    ```
  </details>
