---
title: "R Functions"
subtitle: "Solution Exercises"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 3
    collapsed: false
    css: ["assets/css/custom_report.css"]
linestretch: 1.5
number_sections: true
link-citations: true 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```


# Functions


1. Define a function to transform temperature from Fahrenheit to Celsius
    $$
    Fahrenheit = Celsius *1.8 + 32
    $$
    
    <details><summary>See solution</summary>
    ```{r}
    fahr2cel <- function(x){
      res <- (x - 32) / 1.8
      
      return(res)
    }
    
    fahr2cel(68)
    ```
    </details>
  
1. Define a function `n_and_media()` that given a numeric vector returns the number of elements and the vector mean in a nicely formatted sentence
    > `The length is...`<br>`The mean is  ...`
  
    <details><summary>See solution</summary>
    ```{r}
    n_and_media <- function(x){
      x_length <- length(x)
      x_mean <- mean(x)
      
      cat("The length is", x_length,
          "\nThe mean is", round(x_mean, 3))
    }
    
    n_and_media(1:10)
    ```
    </details>
    
1. Define a function that compute the product between two values. Values need to be interactively defined by the user using `readline()`.

    <details><summary>See solution</summary>
    ```{r}
    my_product <- function(){
      x <- readline(prompt = "First value: ")
      y <- readline(prompt = "Second value: ")
      
      res <- as.numeric(x) * as.numeric(y)
      
      return(res)
    }
    ```
    </details>

# Conditional Programming

1. Create a function that indicate whether a number is even or odd.
    <details><summary>See solution</summary>
    ```{r}
    even_odd <- function(x){
      if(x %% 2 == 0){
        cat(x, "is even")
      } else {
        cat(x, "is odd")
      }
    }
    
    even_odd(4)
    even_odd(5)
    ```
    </details>
    
1. Define a function that assign a mark according to percentage of correct answers
  > F $<$ .55 <br>
  > .55 $<=$  E $<$ .65 <br>
  > .65 $<=$ D $<$ .75  <br>
  > .75 $<=$ C $<$ .85 <br>
  > .85 $<=$ B $<$ .95 <br>
  > .95 $<=$ A
  
    <details><summary>See solution</summary>
    ```{r}
    test_mark <- function(x){
      if(x < .55){
        res <- "F"
      } else if (x < .65) {
        res <- "E"
      } else if (x < .75) {
        res <- "D"
      } else if (x < .85) {
        res <- "C"
      } else if (x < .95) {
        res <- "B"
      } else {
        res <- "A"
      }
      
      return(res)
    }
    
    test_mark(.3)
    test_mark(.7)
    test_mark(.9)
    ```
    </details>


# Iterative Programming

1. Crete a function that computes the mean of a numeric vector.
    <details><summary>See solution</summary>
    ```{r}
    my_mean <- function(x){
      x_length <- length(x)
      
      # initialize values
      x_sum <- 0
      
      # for loop
      for(i in seq_along(x)){
        x_sum <-  x_sum + x[i]
      }
      
      res <- x_sum/x_length
      
      return(res)
    }
    
    my_mean(c(4, 2, -3, 6, 0))
    mean(c(4, 2, -3, 6, 0))
    ```
    </details>
1. Crete a function that returns the min and the max of a numeric vector.
    <details><summary>See solution</summary>
    ```{r}
    my_range <- function(x){
      x_length <- length(x)
      
      # initialize values
      x_min <- x[1]
      x_max <- x[1]
      
      # for loop
      for(i in seq_along(x)[-1]){
        
        # check min
        if(x[i] < x_min){
          x_min <- x[i]
        }
        
        # check max
        if(x[i] > x_max){
          x_max <- x[i]
        }
      }
      
      res <- c(x_min, x_max)
      
      return(res)
    }
    
    my_range(c(4, 2, -3, 6, 0))
    range(c(4, 2, -3, 6, 0))
    ```
    </details>
1. Crete a function that computes the median of a numeric vector.
    <details><summary>See solution</summary>
    ```{r}
    my_median <- function(x){
      x_length <- length(x)
      
      x <- sort(x)
      
      if(x_length %% 2 == 1){
        res <- x[x_length %/% 2 + 1]
      } else {
        res <- mean(c(x[x_length %/% 2], x[x_length %/% 2 + 1]))
      }
      
      return(res)
    }
    
    my_median(x = c(4, 2, 6, 0))
    median(c(4, 2, 6, 0))    
    ```
    </details>

