---
title: "01 Introduction"
subtitle: "Elements of R e R Markdown"
author: "Claudio Zandonella"
institute: ""
output:
  xaringan::moon_reader:
    css: ["assets/custom-presentation.css", "default-fonts", "default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("tidyverse")
library("kableExtra")

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center', dpi = 500,
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      tidy=TRUE, tidy.opts = list(width.cutoff = 80),
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))


options(htmltool.dir.version = FALSE)
xaringanExtra:::use_freezeframe()

tryCatch(
    devtools::load_all("../../"),
    error = function(x){
      devtools::load_all()
      })
```


# The Course

--

<br>

Aim:

> Learning R and R Markdown

---

# The Course

<br>

Aim:

> Learning R and R Markdown... at least we try

--

<br>

Materials and Infos:

> https://arca-dpss.github.io/course-R-Rmarkdown/

---

# Why R?

--

- open science (reproducibility)
- software open-source

--

```{r, fig.width= 5, fig.asp=.6, dpi = 300, out.width="60%"}

colors <- c("SPSS" = "#ED174B", "R" = "#4AA4DE")


ggplot() +
  geom_function(aes(color = "SPSS"),
                fun = function(x) log((x + .1)*10, base = 5) *20,
                size = 2, linetype = "longdash") +
  geom_function(aes(color = "R"),
                fun = function(x) 3^(x), size = 2) +
  xlim(0, 4.5) +
  labs(x = "Learning Time",
       y = "What we can do",
       color = NULL) +
  scale_color_manual(values = colors) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "top",
        legend.background = element_rect(fill="transparent"))
```

---
# Why ~~R~~ Programming?


<br>

--

- It's like magic
--

- We are leaving in 2022 (info-sphere)
--

- Let computers do the boring stuff
--

- Programmers are lazy
--

- Good for OCD and Dopamine circuit
--

<br>

<br>

> We are not learning R.....

---

# Why ~~R~~ Programming?

<br>

- It's like magic
- We are leaving in 2022 (info-sphere)
- Let computers do the boring stuff
- Programmers are lazy
- Good for OCD and Dopamine circuit

<br>

> We are not learning R.....we are learning how to programme

---

# What is R? 

--

- It is not a calculator

--

<br>
- It is not a statistical software

--

<br>

> **It is a programming Language**

---

# Programming Language

--

As any language we have:

<br>

--

- Elements
--

- Rules
--

- Communicate
--

- Evolving

---

class: end, center, middle

# Let's staRt

