---
title: "02 R Data Structures"
subtitle: "Elements of R e R Markdown"
author: "Claudio Zandonella"
institute: ""
output:
  xaringan::moon_reader:
    css: ["assets/custom-presentation.css", "default-fonts", "default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("tidyverse")
library("kableExtra")

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center', dpi = 500,
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))


options(htmltool.dir.version = FALSE)
xaringanExtra:::use_freezeframe()

tryCatch(
    devtools::load_all("../../"),
    error = function(x){
      devtools::load_all()
      })
```
class: size-small
# Math Operators

```{r}
names_function <- c("Addition", "Subtraction", "Multiplication", "Division",
                    "Remainder from division", "Integer Division", "Power",
                    "Absolute value","Sign", "Square Root", 
                    "Logarithm (natural)", "Exponential", "Trigonometriche",
                    "Fattoriale", "Coefficiente binomiale") 

math_operators <- data.frame(
    formula = c(sprintf("x %s y", c("+", "-", "*", "/", "%%", "%/%", "^")),
                sprintf("%s(x)", c("abs", "sign", "sqrt", "log", "exp")),
                paste0(sprintf("%s(x)", c("sin", "cos", "tan", "asin", "acos", "atan")),
                       collapse = "<br />"),
                "factorial(x)", "choose(n, k)"),
    name = names_function,
    example = c(sprintf("> %s <br />[1] %s",
                        c("5 + 3", "7 - 2", "4 * 3", "8 / 3", "7 %% 5", "7 %/% 5", 
                          "3 ^ 3","abs(3-5^2)", "sign(-8)", "sqrt(25)", "log(10)", "exp(1)"),
                        c("8", "5", "12", "2.666667", "2", "1", "27","22", "-1", "5",
                          "2.302585", "2.718282")),
                ">sin(pi/2) <br />[1]1 <br />>cos(pi/2) <br />[1]6.123234e-17",
                c(sprintf("> %s <br />[1] %s",
                          c("factorial(6)", "choose(5,3)"),
                          c("720", "10"))))
  )
```

.pull-left-50[
```{r}
kable(math_operators[1:7, ], col.names = c("Function", "Name", "Example"), escape = FALSE, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
]

.pull-right-50[
```{r}
kable(math_operators[8:12, ], col.names = c("Function", "Name", "Example"), escape = FALSE, format = "html", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
]

---
class: size-small
# Relational and Logical Operators

.pull-left-50[
```{r}
names_function <- c("Equal to", "Not equal to", "Greater", "Greater or equal", 
                    "Less", "Less or equal", "Inclusion") 


logical_operators <- data.frame(
  formula = sprintf("x %s y", c("==", "!=", ">", ">=", "<", "<=", "%in%")),
  name = names_function,
  example = sprintf("> %s <br />[1] %s",
                    c("5 == 3", "7 != 2", "4 > 3", "-2 >= 3", "7 < 5", "7 <= 7", "5 %in% c(3, 5, 8)"),
                    c("FALSE", "TRUE", "TRUE", "FALSE", "FALSE", "TRUE", "TRUE")))


kableExtra::kable(logical_operators, col.names = c("Function", "Name", "Example"),
                  escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE)
```
]

.pull-right-50[
```{r}

names_function <- c("And", "Or (inclusive)", "Or (disgiuntive)", "Negation") 

logical_operators <- data.frame(
  formula = c(sprintf("x %s y", c("&", "|")), "xor(x, y)", "!x"),
  name = names_function,
  example = c(sprintf("> %s <br />[1] %s<br />> %s <br />[1] %s",
                    c("TRUE & TRUE", "TRUE & FALSE", "xor(TRUE, FALSE)"),
                    c("TRUE", "FALSE", "TRUE"),
                    c("TRUE & FALSE", "FALSE & FALSE", "xor(TRUE, TRUE)"),
                    c("FALSE", "FALSE", "FALSE")), "> !TRUE <br />[1] FALSE"))


kableExtra::kable(logical_operators, col.names = c("Function", "Name", "Example"),
                  escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE)
```

]


---
class: size-small

# Operators Order

.pull-left-50[
1. Math operators<br>(e.g., `^`, `*`, `/`, `+`, `-`, etc.)
2. Relational operators<br>(e.g., `<`, `>`, `<=`, `>=`, `==`, `!=`)
3. Logical operators<br>(e.g., `!`, `&`, `|`)
]

.pull-right-50[
1. `^` (Power)
2. `%%` (remainder from division) and<br>`%/%` (integer division) 
3. `*` (multiplication) and<br>`/`(division)
4. `+` (addition) and<br>`-`(subraction)
]

<div style="padding-top:250pt;">

<center><blockquote style ="width: fit-content;">Use parenthesis "`(...)`"!</blockquote></center>

</div>

---
class: size-small
# Vectors


.pull-left-50[
### Creating

```{r, eval=FALSE, echo=TRUE}
name_vector <- c(...)
```
```{r, echo=TRUE}
x <- c(2, 5, 1, 3, 7)
x

# Create a sequence with '<from>:<to>'
y <- 1:8
y
```
]

.pull-right-50[
### Selecting
```{r, eval=FALSE, echo=TRUE}
name_vector[<index-position>]
```


```{r, echo = TRUE}
# Position index
x[c(2, 5, 1)]

# Logical condition
x[x >= 5]

# Removing 
x[- c(1, 3)]

# Editing
x[1] <- 9999
x
```
]

---
class: size-small
# Vectors

.pull-left-50[
### Types

```{r, echo=TRUE}
# Character
c("a", "b", "c")

# Double
c(3.5, 2, 9.34)

# Integer
c(1L, 2L, 3L)

# Logical
c(TRUE, FALSE, FALSE)
```
]

.pull-right-50[
### Function Families

- `is.*()` test if vector of the correct type
- `as.*()` coerces object to given type

### Special Values

- `NULL`, the null object of length 0 (**important** `is.null()`)
- `NA`, logical of length 1 indicating a missing value (**important** `is.na()`)
- `Inf` (`-Inf`), numeric value indicating too large (or too small values)
- `NaN`, numeric value indicating a number that can not be represented

]
---
class: size-small

# Common Functions

.pull-left-50[
- `length()`, get object length
    ```{r, echo = TRUE}
    length(c("a", "b", "c"))
    ```
- `seq()`, sequence generation
    ```{r, echo = TRUE}
    seq(from = 0, to = 10, by = 2)
    ```
- `rep()`, replicate elements
    ```{r, echo = TRUE}
    rep(c("a", "b", "c"), each = 2)
    ```
]

.pull-right-50[
- `sort()`, sort vector values
    ```{r, echo = TRUE}
    sort(c(6,3,8,1,3))
    ```
- `order()`, return permutation which rearranges
    ```{r, echo = TRUE}
    order(c(6,3,8,1,3))
    ```
- `order()`, get position of `TRUE` values
    ```{r, echo = TRUE}
    which(c(6,3,8,1,3) == 3)
    ```
]

---
class: size-small
# Factors

.pull-left-50[
### Creating

```{r, eval=FALSE, echo=TRUE}
factor(x = , levels = , ordered = FALSE)
```
```{r, echo=TRUE}
# Unordered factor
x <- factor(c("a", "b", "a", "c"))
x

# Ordered factor, see also ?ordered()
y <- factor(
  c("IV", "II", "I", "III"),
  levels = c("I", "II", "III", "IV"),
  ordered = TRUE)
y
```
]
.pull-right-50[
### Redefining Levels  
```{r, echo=TRUE}
levels(x)

levels(x) <- c("A", "B", "C")
x
```

### Underline Structure
```{r, echo = TRUE}
str(x)
as.numeric(x)
```
]

---
class: size-small
# Matrices

.pull-left-50[
### Creating

```{r, eval=FALSE, echo=TRUE}
matrix(data, nrow = , ncol = , byrow = FALSE)
```
```{r, echo=TRUE}
X <- matrix(1:12, ncol = 4)
X
```

### Dimension
```{r, echo=TRUE}
nrow(X)
ncol(X)
dim(X)
```
]
.pull-right-50[

### Selecting

```{r, eval=FALSE, echo=TRUE}
name_matrix[<row-index>, <col-index>]
```

```{r, echo = TRUE}
# Single element
X[2, 3]

# Entire row/column
X[3, ]
X[ , 2]

# Subsetting 
X[1:2, 3:4]
```

]

---
class: size-small
# Matrices

.pull-left-50[
### Rows and Cols Names  

```{r, echo=TRUE}
# Assigning  names
colnames(X) <- c("A", "B", "C", "D")
rownames(X) <- c("a", "b", "c")

X

# Selecting by names
X[ , "A"]
```
]

.pull-right-50[
### Operations

```{r}
names_function <- c("Bind matrices by rows",
                    "Bind matrices by columns", 
                    "Matrix transpose",
                    "Get diagonal elements",
                    "Determinant matrix",
                    "Inverse matrix",
                    "Elementwise sum",
                    "Elementwise difference",
                    "Elementwise product",
                    "Elementwise division",
                    "Matrix product")
formula <- sprintf("X %s Y", c("+", "-", "*", "/"))


matrix_operators <- data.frame(formula = c(
  sprintf("%s(X, Y)", c("cbind", "rbind")),
  sprintf("%s(X)", c("t", "diag", "det", "solve")),
  formula, "X %*% Y"),
  name = names_function)

kableExtra::kable(matrix_operators, col.names = c("Functions", "Description"),
                  escape = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE)
```
]
---
class: size-small
# Dataframe

.pull-left-50[
### Creating

```{r, eval=FALSE, echo=TRUE}
data.frame(<var-1> = ...,
           <var-2> = ...,
           stringsAsFactors = FALSE)

```
```{r, echo=TRUE}
df <-  data.frame(
  name = c("A", "B", "C", "D"),
  x = factor(c(0,1,0, 1)),
  y = 1:4)

df
```

]
.pull-right-50[
### Structure

```{r, echo=TRUE}
str(df)
```

### Summary
```{r, echo=TRUE}
summary(df)
```

]

---
class: size-small
# Dataframe

.pull-left-50[
### Selecting

```{r, eval=FALSE, echo=TRUE}
name_df[<row-index>, <col-index>]
```

```{r, echo = TRUE}
# Single value
df[2, 1]

# Entire rows/columns
df[3:4, ]

df[ , 2:3]
```
]

.pull-right-50[

<br>

```{r, echo = TRUE}
# By colnames
df[, c("name", "y")]

df$name

# Logical conditions
df[df$y > 1, c("name", "y")]
```
]
---
class: size-small
# Dataframe

.pull-left-50[
### Rows and Cols Names  

```{r, echo=TRUE}
# Assigning  names
colnames(df) <- c("NAME", "X", "Y")
rownames(df) <- c("a", "b", "c", "d")

df
```
]
.pull-right-50[
### Operations
```{r}
names_function <- c("Number of rows",
                    "Number of columns",
                    "Create new column",
                    "Bind by columns",
                    "Bind by rows", 
                    "Return first rows",
                    "Return last rows",
                    "Struttura del dataframe",
                    "Summary del dataframe")

formula_names <-c("nrow", "ncol", "head", "tail", "str", "summary")

df_operators <- data.frame(formula = c(
  sprintf("%s(df)", formula_names[1:2]),
  "df$nome_var &lt;- var_values",
  "cbind(df, dati)",
  "rbind(df, dati)",
  sprintf("%s(df)", formula_names[3:6])),
  name = names_function)

kableExtra::kable(df_operators, col.names = c("Function", "Description"),
                  escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE)
```
]

---
class: size-small
# List

.pull-left-50[
### Creating

```{r, eval=FALSE, echo=TRUE}
list(<item-1> = ...,
     <item-2> = ...,
     ...)

```
```{r, echo=TRUE}
my_list <-  list(my_vector = x,
                 my_matrix = X,
                 my_df = df)

# Number of items
length(my_list)

# Item names
names(my_list)
```

]
.pull-right-50[
### Structure

```{r, echo=TRUE}
str(my_list, max.level = 2)
```

### Add element
```{r, echo=TRUE, eval= FALSE}
my_list$new_item <- "Hello World!"

c(my_list, new_item = "Hello World!")
```

]

---
class: size-small
# List

.pull-left-50[
### Selecting (as list)

```{r, eval=FALSE, echo=TRUE}
list_name[<name or position-index>]
```
```{r, echo=TRUE}
my_list[1]

my_list["my_vector"]

class(my_list[1])
```

]
.pull-right-50[
### Selecting (extract elementk)

```{r, eval=FALSE, echo=TRUE}
list_name[[<name or position-index>]]

list_name$<name>
```
```{r, echo=TRUE}
my_list[[1]]

my_list[["my_vector"]]

class(my_list$my_vector)
```

]


---
class: end, center, middle

# Thanks!
