---
title: "03 Functions"
subtitle: "Elements of R e R Markdown"
author: "Claudio Zandonella"
institute: ""
output:
  xaringan::moon_reader:
    css: ["assets/custom-presentation.css", "default-fonts", "default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("tidyverse")
library("kableExtra")

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center', dpi = 500,
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))


options(htmltool.dir.version = FALSE)
xaringanExtra:::use_freezeframe()

tryCatch(
    devtools::load_all("../../"),
    error = function(x){
      devtools::load_all()
      })
```
class: size-small
# Defining Functions

.pull-left-50[
```{r, echo = TRUE, eval = FALSE}
name_function <- function(<argument-1>, 
                          <argument-2>, 
                          ...){
  
  <body-function>
    
  return(<result>)
}
```


```{r, echo = TRUE}
greetings <- function(name){
  
  res <- paste("Hello", name)
  
  return(res)
}

greetings("Claudio")
```

]

.pull-right-50[

### Arguments

- `function(x){...}`, argument `x` is require with not default value
- `function(x = 0){...}`, argument `x` is require with default value `0`
- `function(...){...}`, argument `...` allows arbitrary number of arguments

### Interactive

- `cat()` and `print()`, print objects in console
- `readline(prompt = "...")`, get user's response
- `menu(<choices>)`, user select answer from menu

]

---
class: size-small
# Conditional Programming

.pull-left-50[
### If statement

```{r, echo = TRUE, eval = FALSE}
if (<test>) {
  <body>
}
```

```{r, echo = TRUE}
is_negative <- function(x){
  
  if (x < 0) {
    print(paste(x, "is negative"))
  }
}

is_negative(-2)
```
]

.pull-right-50[
### If else statement

```{r, echo = TRUE, eval = FALSE}
if (<test>) {
  <body>
} else {
  <body>
}
```

```{r, echo = TRUE}
tell_sign <- function(x){
  
  if (x < 0) {
    print(paste(x, "is negative"))
  } else {
    print(paste(x, "is possitive"))
  }
}

is_negative(2)
```
]

---
class: size-small
# Iterative Programming

.pull-left-50[
### For loop

```{r, echo = TRUE, eval = FALSE}
for (i in c(...)) {
    <body>
}
```

```{r, echo = TRUE}
call_friends <- function(friends){
  for( friend in friends){
    cat("Hello ", friend, "!\n", sep = "")
  }
}

call_friends(
  friends = c("Emy", "John", "Mat")
)
```
]

.pull-right-50[
### While loop

```{r, echo = TRUE, eval = FALSE}
while (<test>) {
    <body>
}
```

```{r, echo = TRUE}
countdown <- function(count){
  
  while(count > 0){
    cat(count, "\n")
    count <- count - 1
  }
  cat("Start!\n")
}

countdown(count = 3)
```
]

---
class: size-small
# Apply Family

.pull-left-50[
```{r, echo = TRUE, eval = FALSE}
apply(X = , MARGIN = , FUN = )
```

```{r, echo = TRUE}
(X <- matrix(1:12, ncol = 4))

# By rows
apply(X, MARGIN = 1, FUN = sum)

# By columns
apply(X, MARGIN = 2, FUN = sum)

# Custom function
apply(X, MARGIN = 1, function(x){
        round(sd(x)/mean(x), 2) 
      })
```
]

.pull-right-50[
### Other

- `tapply()`, apply a function to each group of values given by a combinations of factors
- `lapply()`, apply a function over all elements of a lists (return a list)
- `sapply()`, apply a function over all elements of a lists (return a vector or a list)
- `vapply()`, apply a function over all elements of a lists (return a vector)
- `replicate()`, repeatedly apply a function.
]

---
class: size-small
# Probability Distributions

.pull-left-50[
### Normal Disribution

- `dnorm(x, mean = 0, sd = 1)`, get the density given the quantile 
- `pnorm(q, mean = 0, sd = 1)`, get the probability given the quantile
- `qnorm(p, mean = 0, sd = 1)`, get the quantile given the probability 
- `rnorm(n, mean = 0, sd = 1)`, sample `n` values 

```{r, echo = TRUE}
dnorm(0, mean = 0, sd = 1)
pnorm(0, mean = 0, sd = 1)
qnorm(0, mean = 0, sd = 1)
rnorm(0, mean = 0, sd = 1)
```
]

.pull-right-50[
### Other Distributions

```{r}
dist_fun <- sprintf('<code class="remark-inline-code">*%s()</code>', 
                    c("binom", "pois", "unif", "chisq", "t", "f", "gamma"))

dist_name <- sprintf('%s', 
                    c("Binomial", "Poisson", "Uniform", "Chi-Square", 
                      "Student-t", "F", "Gamma"))
  
data_table <- data.frame(
  discrete = c(paste0(dist_fun, ", ", dist_name)[1:2], rep(" ", 3)),
  continuous = paste0(dist_fun, ", ", dist_name)[3:7]
)

kable(data_table, col.names = c("Discrete", "Continuous"), escape = FALSE, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Randomness

```{r, echo = TRUE, eval = FALSE}
# Random Number Generation
set.seed(seed = <value>)

# Smaple and permutations
sample(x = , size = , replace = , prob = )

```
]


---
class: end, middle, center

# Thanks!
