---
title: "04 Data-Analysis"
subtitle: "Elements of R e R Markdown"
author: "Claudio Zandonella"
institute: ""
output:
  xaringan::moon_reader:
    css: ["assets/custom-presentation.css", "default-fonts", "default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("tidyverse")
library("kableExtra")

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center', dpi = 500,
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))


options(htmltool.dir.version = FALSE)
xaringanExtra:::use_freezeframe()

tryCatch(
    devtools::load_all("../../"),
    error = function(x){
      devtools::load_all()
      })
```
class: size-small
# Working Session

.pull-left-50[

### Environment

- `ls()`, list all objects in the environment 
- `rm(<name_object>)`, remove object from the environment 
- `rm(list = ls())`, remove all objects from the environment

> Use RStudio Projects!

### Libraries

- `install.packages("<package-name>")`, install package
- `library("<package-name")`, load package
- `<package_name>::<function>()`, use function without loading package

]

.pull-right-50[

### Working Directory

- `getwd()`, get absolute filepath
- `setwd()`, set the working directory in the session

### Paths

- Favour **Relative path** instead of **Absolute path**.
- Use `"/"` (forward slash) to indicate paths
- `"./"` current working directory
- `"../"` parent directory

### Help!....I need somebody, Help!

- `?<name_function>()`, get function documentation
- `??<name>`, search for matches in documentation

]

---

class: size-small
# Managing Data

<br>

```{r}
format <- sprintf('<bold><code class="remark-inline-code">.%s()</code></bold>', 
                  c("rda", "rds", "csv", "csv", "txt", "xlsx"))

load <- sprintf('<code class="remark-inline-code">%s</code>', 
                c('load("&lt;data&gt;.rda")', 
                  'readRDS("&lt;data&gt;.rds")',
                  'read.csv("&lt;data&gt;.csv")', 
                  'read.csv2()',
                  'read.table("&lt;data&gt;.txt", sep = )',
                  'readxl::read_xlsx("&lt;data&gt;.xlsx")'))


save <- sprintf('<code class="remark-inline-code">%s</code>', 
                c('save(&lt;object&gt;, file = "&lt;data&gt;.rda")', 
                  'saveRDS(&lt;object&gt;, file = "&lt;data&gt;.rds")',
                  'write.csv(&lt;object&gt;, file = "&lt;data&gt;.csv")', 
                  'write.csv2(&lt;object&gt;, file = "&lt;data&gt;.csv")',
                  'write.table(&lt;object&gt;, file = "&lt;data&gt;.txt")',
                  'really??'))

format[4] <-  " "
load[3] <- paste0(load[3], ' for <code class="remark-inline-code">,</code> as separator and <code class="remark-inline-code">.</code> as decimal separator')
load[4] <- paste0(load[4], ' for <code class="remark-inline-code">;</code> as separator and <code class="remark-inline-code">,</code> as decimal separator')

data_table <- data.frame(
  format = format,
  load = load,
  save = save
)

kable(data_table, col.names = c("Format", "Read", "Write"), escape = FALSE, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  column_spec(2, width = "12cm") %>% 
  column_spec(3, width = "13cm")
```


---
class: size-small

# Descriptive Analysis

<br>

.pull-left-50[
- `summary()`, summary information
- `mean()`, arithmetic mean 
- `median()`, median value
- `quantile()`, sample quantiles
- `max()`, maximum value
- `min()`, minimum value
- `range()`, range of values
- `sd()`, standard deviation 
- `var()`, variance
- `cor()`, correlation
- `cov()`, covariance
- `table()`, frequency table
]

.pull-right-50[

### Utility Functions
- `paste()`, concatenate string
- `paste0()`, concatenate string
- `head()`, return first rows
- `cumsum()`, cumulative sum
- `cumprod()`, cumulative product
- `round()`, rounding numbers
- `scale()`, scaling and centering values
- `unique()`, get unique elements
- `duplicated()`, determine duplicated elements
- `complete.cases()`, find complete cases
]

---
class: size-small
# Plots

.pull-left-50[
- `plot(x, y, ...)`, scatter plot
- `plot(x, y, type = "l",...)`, line plot
- `barplot(height, ...)`, barplot
- `hist(x, ...)`, histogram
- `boxplot(formula, ...)` boxplot
- `plot(density(x, ...))`, kernel density
- `pairs()`, matrices of scatter plots

### Plot Options

- `col`, color of lines and points 
- `main`, main plot title
- `xlab` / `ylab`, label for x/y axis
- `xlim` / `ylim`, limits for x/y axis
- `legend()`, add legend
- `par(mfrow = c(<n-rows>, <n-cols>))`, <br>define plots grid

]

.pull-right-50[
```{r example-plot}
data(mtcars)
x <- 1:10/10
y <-  2 * x + rnorm(10)

par(mfrow = c(3, 2))

plot(mtcars$mpg, mtcars$hp, main = "plot(mtcars$mpg, mtcars$hp)")
plot(x, y, type = "l", main = 'plot(x, y, type = "l")')
barplot(table(mtcars$cyl), xlab = "cyl", main = 'barplot(table(mtcars$cyl), xlab = "cyl")')
boxplot(mpg ~ cyl, data = mtcars, main = 'boxplot(mpg ~ cyl, data = mtcars)')
hist(mtcars$hp, main = 'hist(mtcars$hp)')
plot(density(mtcars$hp), main = 'plot(density(mtcars$hp))')

par(mfrow = c(1, 1))


```

]

---
class: size-small
# Statistics

.pull-left-50[
### Tests
- `cor.test()`, test correlation
- `t.test()`, Student's t-test
- `wilcox.test()`, Wilcoxon rank test
- `fisher.test()`, Fisher's exact test
- `chisq.test()`, Pearson's chi-squared test 
- `shapiro.test()`, normality Test

### Models

- `lm()`, fit linear models
- `glm()`, fit generalized linear models
- `anova()`, compute analysis of variance
]

.pull-right-50[
### Formula ~

`<outcome> ~ <predictors>`

- `1` add intercept term
- `a + b` additive effect
- `a:b` interaction effect
- `a*b` interaction and lower terms
- `(a + b + ...)^n` limit to interaction of grade `n`
- `- a` remove term

### Contrasts

- `contr.treatment()`, dummy coding 
- `contr.sum()`, sum-to-zero contrasts
- `contr.poly()`, orthogonal polynomial contrasts
- `contr.helmert()`, Helmert contrasts
]

---
class: size-small
# Toop Package

.pull-left-50[
#### Models

- `lme4`, fit linear and generalized linear mixed-effects models
- `brms`, fit Bayesian generalized (non-)linear multivariate multilevel models using 'Stan' for full Bayesian inference.
- `metafor`, a comprehensive collection of functions for conducting meta-analyses in R.
- `lavaan`, fit a variety of latent variable models, including confirmatory factor analysis, structural equation modeling and latent growth curve models
]

.pull-right-50[
#### Utilities

- `car`, variety of tests and diagnostics for regresion analysis
- `performance`, utilities for computing measures to assess model quality.
- `emmeans`, obtain estimated marginal means (EMMs) and cntrasts for many linear, generalized linear, and mixed models.



]
---
class: end, middle, center

# Thanks!
