---
title: "Exercises R Data Structure"
subtitle: "Elements of R e R Markdown"
author: "Claudio Zandonella"
institute: ""
output:
  xaringan::moon_reader:
    css: ["assets/custom-presentation.css", "default-fonts", "default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("tidyverse")
library("kableExtra")

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center', dpi = 500,
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

options(crayon.enabled = TRUE)

old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)
# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))


options(htmltool.dir.version = FALSE)
xaringanExtra:::use_freezeframe()

tryCatch(
    devtools::load_all("../../"),
    error = function(x){
      devtools::load_all()
      })
```

class: size-small
# Operators

.pull-left-50[
#### For Coding Pinguins 🐧


1. Compute $\frac{(45+21)^3+\frac{3}{4}}{\sqrt{32-\frac{12}{17}}}$.
1. Compute $\frac{\sqrt{7-\pi}}{3\ (45-34)}$.
1. Define a proposition to evaluate if a number is even.
1. Define a proposition to evaluate if a number is between -4 and -2 or between 2 and 4.
1. Compare the returned results by<br> `4 ^ 3 %in% c(2,3,4)` and <br>`4 * 3 %in% c(2,3,4)`.
]

.pull-right-50[

#### For Coding Monkeys 🐒

1. Which is the difference between `==` and `all.equal()`?
1. Compare the returned results by `TRUE & NA` and `FALSE & NA`
1. Compare the returned results by `TRUE | NA` and `FALSE | NA`
1. Which is the difference between `&` (o `|`) and<br> `&&` (o `||`)?
1. Do you know the functions `isTRUE()` and `isFALSE()`?
]

---
class: size-small
# Vectors

.pull-left-50[
#### For Coding Pinguins 🐧

Creation (`?seq()` and `?rep()`)

1. Create `x` with all even numbers between 1 e 25 .
1. Create `y` with the first 10 multiples of 7 starting from 14.
1. Create `a` with the letters `"A"`,`"B"` e `"C"` repeated in the same order 4 times.
1. Create `b` with the letters `"A"`,`"B"` e `"C"` each one repeated 4 times.

Selection

1. From `x`, select the values 8, 12, and 16.
1. From `y`, select values smaller than 36 or grater than 54.
1. Create `c` from `a` removing all `"B"` elements.
1. Crate `d` from `b` substituting all `"B"` with `"D"`.

]

.pull-right-50[

#### For Coding Monkeys 🐒

1. What do you know about named vectors?
1. Why we have both integer and double vectors?
1. Do you know the function `which()`?
1. Which is the meaning of `mean()` and `sum()` on logical vectors.
1. Do you know the functions `any()` and `all()`?
1. Which data type is `NA`?
1. Check the length of vectors with `NA` or with `NULL` values.
]

---
class: size-small

# Factors

.pull-left-50[
#### For Coding Pinguins 🐧

1. Create the factor `gender` as follows,
    ```{r, echo=FALSE}
    factor(c(rep(c("M","F"),3),"F","F","M"))
    ```
1. Rename levels in `"Males"` and `"Females"`
1. Create the factor `size` as follows,
    ```{r, echo=TRUE, eval = FALSE}
    c("Small", "Large", "Medium", 
      "X-Large", "X-Small")
    ```
1. Make `size` an ordered factor, with appropriate label order.
]

.pull-right-50[
#### For Coding Monkeys 🐒

1. How can you get an object attributes?
1. What do you know about classes?
1. Check the data type of a factor?
1. Why a factor vector is better than a string vector?
1. Create a factor `grade` as follows,
    ```{r, echo=FALSE}
    c(5, 4, 3, 3, 5, 4, 4)
    ```
    and convert back to numeric.
]

---
class: size-small

# Matrices

.pull-left-50[
#### For Coding Pinguins 🐧

#### Creating

1. Create the matrix `A` as follows, 

$$
\begin{matrix}
2 & 34 & 12 & 7\\
46 & 93 & 27 & 99\\
23  & 38 & 7 & 04
\end{matrix}
$$
1. Create the matrix `B`4X3 in which `"A"`,`"B"` e `"C"` are repeated at each row.

#### Selecting

1. From `A`, select the value 27.
1. From `B`, select the elements between the second and forth row, second and third column.
1. Form `A`, select only odd values.
1. From `B`, select all values different from `"C"`. 
]

.pull-right-50[
#### For Coding Monkeys 🐒

1. Create a 5x5 matrix with named columns and rows.
1. Try different selections using names. 
1. A vector is a column vector or a row vector?
1. Selecting a column returns a vector or a matrix?
1. Check the different uses of `diag()`.
1. Do you know the function `array()`?

]

---
class: size-small

# Dataframe

.pull-left-50[
#### For Coding Pinguins 🐧

Creating

1. Create the following dataframe

```{r, echo = FALSE}
data_wide<-data.frame(Id = c("subj_1","subj_2","subj_3"),
                      age = c(21,23,19),
                      gender = c("F","M","F"),
                      item_1 = c(2,1,1),
                      item_2 = c(0,2,1),
                      item_3 = c(2,0,1))
data_wide
```

Selecting

1. Select data of `"subj_3"`.
1. Select only the answers to the items.
1. Select `Id` and `gender` for subject with `1` at `item_1`.
]

.pull-right-50[
#### For Coding Monkeys 🐒

1. Check the options `stringsAsFactors` in the function `data.frame()`.
1. Transform the dataframe from the wide-format to the long-format.
1. Comment the different results obtained by `data_wide$Id`,<br> `class(data_wide[, "Id"])`, and<br> `class(data_wide["Id"])`

]

---
class: size-small

# Lists

.pull-left-50[
#### For Coding Pinguins 🐧

Creating

1. Create a list with the elements form the previous exercises:
  - vector `x`
  - matrix `A`
  - dataframe

Selecting

1. Select the vector and the matrix from the list.
1. Extract the dataframe from the list.
1. Add the sting `"Hello World!"` to the list.
]

.pull-right-50[
#### For Coding Monkeys 🐒

1. Select the age value `23` from the dataframe in the list in  single command. 
1. Crete an unnamed list.
1. Add names to an unnamed list.
1. Comment the different results obtained by `length(c(list(a = 1), b = NULL))` and
`length(list(a = 1, b = NULL))`.

]

---
class: end, middle, center

# Thanks!
